---
title: "Guide projetc - Dataquest"
output:
  word_document: default
  html_document: default
  pdf_document: default
---


```{r Packages, message=FALSE, warning=FALSE}

library(readr)
library(purrr)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(knitr)
library(rmarkdown)

```


### [Data - 2011 NYC School Survey](https://data.cityofnewyork.us/Education/2011-NYC-School-Survey/mnz3-dyi8)


### Guided Project: NYC Schools Perceptions  
  * What the project is about
    + It an analysis about the responses to surveys to avaluate the parents, students, and teachers perceptions of the quality of New York City schools
  * What your goal is in this project.
    + Analyse the data to be able to answer questions about their perceptions
        
## Questions:  
  1. Do student, teacher, and parent perceptions of NYC school quality appear to be related to **demographic** and **academic** success metrics?  
  2. Do students, teachers, and parents have similar perceptions of NYC school quality?
  
  
```{r Dataframes, message=FALSE, warning=FALSE, paged.print=TRUE}

## Here is the data already cleaned 
combined <- read_csv("combined.csv")


## Here are the new data that we going to work with 
master_file_11_d75_final <- read_tsv("masterfile11_d75_final.txt")
master_file_11_gened_final <- read_tsv("masterfile11_gened_final.txt")

``` 

In the survey dictionary we could see the names and the descriptions of the variables in this 2 dataframes, and the first column "dbn" could be use as a key variable to join this 2 dataframes because correspond to the School identification 
 
 
###  Selecting the most important variables  

Survey questions assess the following metrics:

* Safety and Respect
* Communication
* Engagement
* Academic Expectations  
  
The following groups responded to the survey:

* Parents
* Teachers
* Students
* Total (The average of parent, teacher, and student scores)
  
So I selected only the variables that correspond to the aggregating scores based on survey results.


### 1. master_file_11_d75_final

```{r dataframe 1, message=FALSE, warning=FALSE, paged.print=TRUE}

master_file_11_d75_final

dim(master_file_11_d75_final)


master_file_11_d75_final_simplified <- master_file_11_d75_final %>% dplyr::select (1:32) %>% dplyr:: select(- highschool)
master_file_11_d75_final_simplified
names(master_file_11_d75_final_simplified)
dim(master_file_11_d75_final_simplified)


```
 
### 2. master_file_11_gened_final

```{r Dataframe 2, message=FALSE, warning=FALSE, paged.print=TRUE}

master_file_11_gened_final
dim(master_file_11_gened_final)

master_file_11_gened_final_simplified <- master_file_11_gened_final %>% dplyr::select (1:32) %>%   dplyr::select(- highschool)%>%   dplyr::filter(schooltype == "High School" | schooltype == "Elementary / Middle School" | schooltype == "Elementary / Middle / High School"  | schooltype == "Middle / High School")          

names(master_file_11_gened_final_simplified )
dim(master_file_11_gened_final_simplified )
paged_table (master_file_11_gened_final_simplified )

```

###  Joinig the 2 dataframes 

Now that we have the dataframes simplified, we can join them.  

With the function "bind_rows()" we can join dataframes from columns that appears in any of the input dataframes:

```{r Joining dataframes, message=FALSE, warning=FALSE, paged.print=TRUE}

dim(master_file_11_gened_final_simplified)
dim(master_file_11_d75_final_simplified)

unified_dataframes <- dplyr::bind_rows(master_file_11_gened_final_simplified , master_file_11_d75_final_simplified)

dim(unified_dataframes)

```  

Now, to join to the dataframe "combined" we need to change the name of the first column of the "unified_dataframe" so the function "_join ()" can find the key variable. 
  
The "combined" dataframe has information about the SATs average of all schools 

```{r full_join, message=FALSE, warning=FALSE, paged.print=TRUE}

# rename() function
unified_dataframes <- unified_dataframes %>% dplyr::rename(DBN =dbn)

# I use full_join to keep all observations of both dataframes 
dataframe_schools_survey <- combined %>%
  full_join(unified_dataframes, by = "DBN")

names(dataframe_schools_survey)
dim(dataframe_schools_survey)

```


## Now that we've created a single, clean dataframe to work with, we can begin our analysis  

### Our first question:  

#### 1. Do student, teacher, and parent perceptions of NYC school quality appear to be related to demographic and academic success metrics?

```{r correlation matrix, message=FALSE, warning=FALSE, paged.print=TRUE}

# selecting variables to compose the correlation matrix 
cor_mat <- dataframe_schools_survey %>%
  select(avg_sat_score, saf_p_11:aca_tot_11) %>% 
  cor(use = "pairwise.complete.obs")

cor_tib <- cor_mat %>%
  as_tibble(rownames = "variable")
dim(cor_tib)


apscore_cors <- cor_tib %>%
  select(variable, avg_sat_score) %>%
  filter(avg_sat_score > 0.25 | avg_sat_score < -0.25)

paged_table(apscore_cors)

# This variables showed a positive correlation with the SAT average 

``` 


Seems like the metrics "Safety and Respect" and "Academic Expectations" showed a correlation with the SATs average. 

Now we going to see this relationship in graphics
  
  

```{r graphics, message=FALSE, warning=FALSE}
# saf_t_11
ggplot(data = dataframe_schools_survey,
       aes(x = saf_t_11, y = avg_sat_score)) +
  geom_point() +
  labs(title = "Correlation between SATs scores and survey answers", 
       y = "Average SATs score", x= "Safety and Respect score - Teachers") + theme(
         panel.background = element_rect(fill = "white")
       )

#saf_s_11
ggplot(data = dataframe_schools_survey,
       aes(x = saf_s_11, y = avg_sat_score)) +
  geom_point() +
  labs(title = "Correlation between SATs scores and survey answers", 
       y = "Average SATs score", x= "Safety and Respect score - Students") + theme(
         panel.background = element_rect(fill = "white")
       )

#aca_s_11
ggplot(data = dataframe_schools_survey,
       aes(x = aca_s_11, y = avg_sat_score)) +
  geom_point() +
  labs(title = "Correlation between SATs scores and survey answers", 
       y = "Average SATs score", x= "Academic Expectations score - Students") + theme(
         panel.background = element_rect(fill = "white")
       )

#saf_tot_11
ggplot(data = dataframe_schools_survey,
       aes(x = saf_tot_11, y = avg_sat_score)) +
  geom_point() +
  labs(title = "Correlation between SATs scores and survey answers", 
       y = "Average SATs score", x= "Safety and Respect score - total") + theme(
         panel.background = element_rect(fill = "white")
       )



```

### It appears that the higher the metrics score, the higher the SAT average.

### **so YES! Student, teacher, and parent perceptions of NYC school quality appear to be related to demographic and academic success metrics**

Now, for our second Questios:
2. Do students, teachers, and parents have similar perceptions of NYC school quality?

```{r pivor_longer, message=FALSE, warning=FALSE, paged.print=TRUE}

# We going to tranform the dataframe in a longer data frame using pivot_longer () function
combined_data_longer <- dataframe_schools_survey %>%
  pivot_longer(cols = saf_p_11:aca_tot_11, 
               names_to = "response_type", 
               values_to = "score")


combined_survey_gather <- combined_data_longer %>% 
  mutate(response = str_sub(response_type, 5,5)) %>% 
  mutate (response_2 = case_when(response == "p" ~ "Parents", 
                                 response == "s" ~ "Students", 
                                 response == "t"~ "Teacher")) %>%
  mutate (metric = str_sub (response_type, 1,3)) %>% dplyr::rename(survey_question = response_type, responses_type = response_2)



# Now we have separeted columns with information about the response type, the metric and the scores 

combined_survey_gather

```


```{r boxplot, message=FALSE, warning=FALSE}

# To see if has difference between the response types, we can do a boxplot using ggplot 
ggplot(combined_survey_gather, aes(x = reorder(responses_type, score, FUN = mean), y = score))+ geom_boxplot (aes( color = responses_type)) + geom_point(aes(color=responses_type, alpha=0.5))

```


## Our results sugest that, in avarege, parents seems to have a better perception about the schools metrics than students and teachers 

